Blame and Coercion: Together Again for the Second Time

### 2019-10-08 paper.tex
LazyD Blame Calculus CEK

\newtheorem{theorem}{Theorem}[]
\newtheorem{proposition}{Proposition}[]
\newtheorem{corollary}{Corollary}[]
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\stxrule}[3]{#1 & ::= & #3 & \text{#2}\\}
\newcommand{\stxrulecont}[1]{& | & #1 & \\}
\newcommand{\funrule}[3]{#1 & = & #2 & #3\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^\ell & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}
\newcommand{\lazyUD}{$\mathtt{Lazy UD}$}
\newcommand{\lazyD}{$\mathtt{Lazy D}$}
\newcommand{\eager}{$\mathtt{Eager}$}
\newcommand{\hyperCoercionI}[0]{id\star}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}

@@

Around 2006, many work on integrating dynamic typing and static typing emerge, 
notably gradual types \cite{siek2006gradual} and hybrid types 
\cite{flanagan2006hybrid}. 
Later \citet{wadler2009well} introduces Blame Calculus, an intermediate 
language for gradual types and hybrid types. 
Blame Calculus throws a blame label when a runtime type-check fails.
Blame labels can include useful information on the failed cast, for example, 
its location in the source code.

@@

In 2007, \citet{herman2010space} proposes a solution to this problem by 
translating Blame Calculus to a variant of \citet{henglein1994dynamic}'s 
Coercion Calculus. Their key idea is to represent casts with coercions, 
which can be composed and normalized. Every coercion looks like a list of 
trees. For normalized coercion, the length of the list is bounded by a small 
constant. And the depths of trees do not growth in composition and 
normalization. These two facts implies that by representing casts with 
coercions, we can restore the space-efficiency of tail-recursive programs. This 
Coercion Calculus, however, doesn't include blame tracking.

@@

\citet{siek2009exploring} incorporate blame tracking into Coercion Calculus by 
decorating coercions with labels. They also propose that there are four blame 
strategies for their Coercion Calculi:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but also blame more programs.
$ D $ and $ UD $ assign blame labels differently.
Blame strategy is essentially the semantics of casts.
They prove that the \lazyUD{} Coercion Calculus simulates the Blame Calculus 
in \cite{wadler2009well}. But we don't know whether the simulation also works 
the other way, and what are the Blame counterparts of other Coercion Calculi.

\citet{siek2010threesomes} proposes another approach to combine blame 
tracking and space efficiency. Their solution is based on Threesome Calculus. 
Their key idea is to represent casts with threesomes, a novel cast 
representation. Threesomes, like coercions, can be composed. There is no 
separate normalization because every threesome is in normal form. 
They prove that their Threesome Calculus bi-simulates
the \lazyUD{} Coercion Calculus and the Blame Calculus.

\citet{siek2012interpretations} introduces a \lazyD{} Blame Calculus. They 
conjecture that it bi-simulates the \lazyD{} Coercion Calculus.

Following \citet{siek2010threesomes}, \citet{garcia2013calculating} 
implement all other blame strategies with Threesome Calculus. 
He claims that 
coercion with labels is easy to understand but hard to implement, 
and that 
threesome with labels, however, is easy to implement but hard to understand.
His claim is later affirmed by the group of people who develop threesome 
(\citet{siek2015blame}).
The connection 
between his Threesome Calculi and \citet{siek2009exploring}'s Coercion 
Calculi are established by the fact that the former are derived from the latter.
The connection between eager calculi and Blame Calculus, however, 
is still unclear.

\citet{siek2015blame} revisit the coercion-based approach. They simplify the 
\lazyUD{} Coercion Calculus by only working with normalized coercions.
Fortunately, coercions are trivially normalized when initially constructed, and 
their empowered compose function produces normalized coercions as well. Again 
\cite{siek2010threesomes}, they delegate all other blame strategies.

Last year, \citet{kuhlenschmidt2018efficient} present Grift, a space-efficient 
and blame-tracking compiler for a gradually typed language of the same name. 
This implementation is based on the \lazyD{} Coercion Calculus.
Their result suggests that implementing coercion is practical. They implement 
tuples, which is the first time gradual product types is considered. Their 
treatment on tuples, however, is shown incorrect 
\cite{Gradual-TypingType-based-casts}.

@@

So far, it seems that the coercion-based approach is the best way to combine 
space-efficiency and blame-tracking: coercion is claimed easy to 
understand \cite{garcia2013calculating}\cite{siek2015blame} and is shown 
easy enough to implement in a compiler \cite{kuhlenschmidt2018efficient}.
Coercion is not satisfactory, however, in at least two aspects.
Firstly, its compose function is not structurally recursive, which makes 
it difficult to study coercion formally.
Many developers of Grift report that it is tricky to convince their 
proof assistants that composition terminates. 
Secondly, it is desirable to understand the nature of ``normalized cast'', but 
the coercion-based approach capture this concept indirectly because normalized 
coercion is defined ny taking a subset of coercion.

@@

Our hyper-coercion considers sum types and product types, which are not 
accounted in all proofs above. Adding each of them requires us to add only one 
new constructor to a component of hyper-coercion. This suggests that 
hyper-coercion might scale better than super-coercion.

We prove formally that the \lazyD{} (resp. \lazyUD{}) hyper-coercion calculus 
bi-simulate the \lazyD{} (resp. \lazyUD{}) Blame Calculus. This is possibly 
the first bi-simulation proof for the \lazyD{} Blame Calculus.

@@

\begin{figure}
	\[
	\begin{array}{lclr}
	\stxrule{S,T}{types}{
		\star \mid{}
		P
	}
	\stxrule{P,Q}{pre-types}{
		\iota \mid{}
		T_1 \rightarrow T_2 \mid{}
		T_1 \times T_2 \mid{}
		T_1 \plus T_2
	}
%	\stxrule{x,y,z}{variables}{\dots}
%	\stxrule{l}{labels}{\dots}
%	\stxrule{k}{constants}{sole}
	\stxrule{e}{terms}{
		x \mid{}
		\mathtt{tt} \mid{}
		\lambda^{S\rightarrow{}T}x.t \mid{}
		e_1 \; e_2
	}
	\stxrulecont{
		\mathtt{cons}\; e_1 \; e_2 \mid{}
		\mathtt{car}\; e \mid{}
		\mathtt{cdr}\; t
	}
	\stxrulecont{
		\mathtt{inl} \; e \mid{}
		\mathtt{inr} \; e \mid{}
		\mathtt{case}\; e_1 \; e_2 \; e_3
	}
	\stxrulecont{
		\langle T \Leftarrow^l S \rangle t \mid{}
		\mathtt{blame} \; l
	}
	\stxrule{o}{observations}{
		\mathtt{tt} \mid{}
		\mathtt{fun}\mid{}
		\mathtt{cons}\mid{}
		\mathtt{inl}\mid{}
		\mathtt{inr}\mid{}
		\mathtt{blame}\; l
	}
	\end{array}
	\]

	\fbox{$ S \sim T $}
	\begin{gather*}
	\inference{}{
		\star \sim \star
	} \quad
	\inference{}{
		\star \sim P
	} \quad
	\inference{}{
		P \sim \star
	} \\
	\inference{}{
		\iota \sim \iota
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \rightarrow T_1 \sim S_2 \rightarrow T_2
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \times T_1 \sim S_2 \times T_2
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \plus T_1 \sim S_2 \plus T_2
	}
	\end{gather*}

	\fbox{$ S \smile T $}
	\begin{gather*}
	\inference{}{
		\star \smile \star
	} \quad
	\inference{}{
		\star \smile P
	} \quad
	\inference{}{
		P \smile \star
	} \\
	\inference{}{
		\iota \smile \iota
	} \quad
	\inference{}{
		S_1 \rightarrow T_1 \smile S_2 \rightarrow T_2
	}
	\inference{}{
	S_1 \times T_1 \smile S_2 \times T_2
	}
	\inference{}{
	S_1 \plus T_1 \smile S_2 \plus T_2
	}
	\end{gather*}

	\fbox{$ \Gamma \vdash e \vdash T $}
	\begin{gather*}
		\inference{
			S \sim T & \Gamma \vdash e : S 
		}{
			\judgetype{\Gamma}{\langle T \Leftarrow^l S \rangle t}{T}
		} \quad
		\inference{
		}{
			\judgetype{\Gamma}{\mathtt{blame} \; l}{T}
		}
	\end{gather*}

@@

$ S \sim T $ reads ``$ S $ and $ T $ are consistent''.
Two types are consistent if one of them is $ \star $, or they have the same 
top-most type constructor and the corresponding sub-parts are consistent. 
The intuition of $ S \sim T $ is that $ S $ and $ T $ have no conflict type 
information. Consistency is reflexive and symmetric, but not transitive.

One might be tempted to conclude that inconsistency is the root of all runtime 
type errors (blames). In the setting of gradual typing, however, runtime 
type-checking is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
top-most type constructor. Shallow-inconsistency is the root of all blames in 
all lazy strategies -- casting a value to a shallowly inconsistent type leads 
to a blame immediately. Shallow-consistency is also reflexive, symmetric, and 
not transitive.

@@

All existing proofs describe the semantics of \lazyD{} Blame Calculus in the 
style of reduction semantics. 
We instead do it with the CEK machine (\citet{felleisen1986control}). 
We choose CEK machine because to achieve space efficiency, we need to change 
continuations. Thus, abstract machines that include continuation in states is 
quite convenient for us. And CEK is one of the simplest machine of this kind.

@@

	\fbox{$ m \fatsemi^\ell (m,t) = (m,t) $}
	\[ 
	\begin{array}{rclclr}
	\comprule{U}{(U,t)}{
		(U,t)
	}{}
	\comprule{c_1 \rightarrow c_2}{(c_3 \rightarrow c_4,t)}{
		(c_3 \fatsemi^{l} c_1 \rightarrow c_2 \fatsemi^\ell c_4, t)
	}{}
	\comprule{c_1 \times c_2}{(c_3 \times c_4,t)}{
		(c_1 \fatsemi^{l} c_3 \times c_2 \fatsemi^\ell c_4, t)
	}{}
	\comprule{c_1 \plus c_2}{(c_3 \plus c_4,t)}{
		(c_1 \fatsemi^{l} c_3 \plus c_2 \fatsemi^\ell c_4, t)
	}{}
	\comprule{b_1}{(b_2,t)}{
		(b_1,\bot^l)
	}{
		\ell = l \; \text{and} \;
		\neg b_1 \smile b_2 
	}
	\end{array}
	\]

@@

	\fbox{$ id( P ) = m $}
	\[
	\begin{array}{rclr}
	\funrule{id(U)}{U}{}
	\funrule{id(T_1 \rightarrow T_2)}{
		id(T_1) \rightarrow id(T_2)
	}{}
	\funrule{id(T_1 \times T_2)}{
		id(T_1) \times id(T_2)
	}{}
	\funrule{id(T_1 \plus T_2)}{
		id(T_1) \plus id(T_2)
	}{}
	\end{array}
	\]

>>>>

\newtheorem{theorem}{Theorem}[]
\newtheorem{proposition}{Proposition}[]
\newtheorem{corollary}{Corollary}[]
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\stxrule}[3]{#1 & ::= & #3 & \text{#2}\\}
\newcommand{\stxrulecont}[1]{& | & #1 & \\}
\newcommand{\funrule}[3]{#1 &=& #2 & #3\\}
\newcommand{\redrule}[3]{#1 & \longrightarrow_B & #2 & #3\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^\ell & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}
\newcommand{\lazyUD}{$\mathtt{Lazy UD}$}
\newcommand{\lazyD}{$\mathtt{Lazy D}$}
\newcommand{\eager}{$\mathtt{Eager}$}
\newcommand{\hyperCoercionI}[0]{id\star}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}
\newcommand{\sOOinspect}[3]{\mathtt{Eval} \; #1 \; #2 \; #3}
\newcommand{\sOOreturn}[2]{\mathtt{Apply} \; #2 \; #1}
\newcommand{\sOOhalt}[1]{\mathtt{Halt} \; #1}
\newcommand{\TOOdyn}[0]{\star}
\newcommand{\TOOpre}[1]{#1}
\newcommand{\POOunit}[0]{\iota}
\newcommand{\POOfun}[2]{#1 \shortrightarrow #2}
\newcommand{\POOprod}[2]{#1 \times #2}
\newcommand{\POOsum}[2]{#1 \plus #2}
\newcommand{\eOOvar}[1]{#1}
\newcommand{\eOOsole}[0]{\mathtt{tt}}
\newcommand{\eOOlam}[4]{\lambda^{#1\rightarrow{}#2}#3.#4}
\newcommand{\eOOapp}[2]{#1 \; #2}
\newcommand{\eOOcons}[2]{\mathtt{cons} \; #1 \; #2}
\newcommand{\eOOcar}[1]{\mathtt{car} \; #1}
\newcommand{\eOOcdr}[1]{\mathtt{cdr} \; #1}
\newcommand{\eOOinl}[1]{\mathtt{inl} \; #1}
\newcommand{\eOOinr}[1]{\mathtt{inr} \; #1}
\newcommand{\eOOcase}[3]{\mathtt{case} \; #1 \; #2 \; #3}
\newcommand{\eOOcast}[4]{#1 \langle \cOOcast{#2}{#3}{#4} \rangle}
\newcommand{\eOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\oOOinj}{\mathtt{dyn}}
\newcommand{\oOOsole}{\mathtt{tt}}
\newcommand{\oOOfun}{\mathtt{fun}}
\newcommand{\oOOcons}{\mathtt{cons}}
\newcommand{\oOOinl}{\mathtt{inl}}
\newcommand{\oOOinr}{\mathtt{inr}}
\newcommand{\oOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\cOOcast}[3]{#1 \overset{#2}{\Rightarrow} #3}
\newcommand{\vOOcast}[2]{#1\langle#2\rangle}
\newcommand{\vOOfun}[3]{\mathtt{fun} \; #1 \; #2 \; #3}
\newcommand{\vOOtt}[0]{\mathtt{tt}}
\newcommand{\vOOcons}[2]{\mathtt{cons}\;#1\;#2}
\newcommand{\vOOinl}[1]{\mathtt{inl}\;#1}
\newcommand{\vOOinr}[1]{\mathtt{inr}\;#1}
\newcommand{\rOOsucc}[1]{\mathtt{succ}\;#1}
\newcommand{\rOOfail}[1]{\mathtt{fail}\;#1}

@@

Around 2006, many work on integrating dynamic typing and static typing emerge, 
notably gradual types \cite{siek2006gradual} and hybrid types 
\cite{flanagan2006hybrid}. 
Later \citet{wadler2009well} introduces Blame Calculus, an intermediate 
language for gradual types and hybrid types. 
Blame Calculus throws a blame label when a runtime type-check fails.
These labels can include useful information on the failed cast, for example, 
its location in the source code.

@@

In 2007, \citet{herman2010space} proposes a solution to this problem by 
translating Blame Calculus to a variant of \citet{henglein1994dynamic}'s 
Coercion Calculus. Their key idea is to represent casts with coercions, 
which can be composed and normalized. Every coercion looks like a list of 
trees. For normalized coercion, the length of the list is bounded by a small 
constant. And the depths of trees do not grow in composition and 
normalization. These two facts implies that by representing casts with 
coercions, we can restore the space-efficiency. This 
Coercion Calculus, however, does not support blame tracking.

@@

\citet{siek2009exploring} incorporate blame tracking into Coercion Calculus by 
decorating coercions with labels. They also propose that there are four blame 
strategies for their Coercion Calculi:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but also blame more programs.
$ D $ and $ UD $ assign blame labels differently.
Blame strategy is essentially the semantics of casts.
They prove that the \lazyUD{} Coercion Calculus simulates the Blame Calculus 
in \cite{wadler2009well}. But it is unknown whether the simulation also works 
the other way, and what the Blame counterparts of other Coercion Calculi are.

\citet{siek2010threesomes} proposes another approach to combine blame 
tracking and space efficiency. Their solution is based on Threesome Calculus. 
Their key idea is to represent casts with threesomes, a novel cast 
representation. Threesomes, like coercions, can be composed. There is no 
separate normalization because every threesome is in normal form. 
They prove that their Threesome Calculus bi-simulates
the \lazyUD{} Coercion Calculus and the Blame Calculus, and that threesomes are 
space-efficient.

\citet{siek2012interpretations} present a \lazyD{} Blame Calculus and revised 
lazy Coercion Calculi. They simplify the Coercion Calculi by only working with 
normal coercions (i.e. coercions in normal form). 
Fortunately, coercions are trivially normal when initially constructed, and 
their empowered compose function is conjectured to produces normal coercions as 
well. Also, they conjecture that their Coercion Calculi bi-simulate the 
corresponding Blame Calculi.

Following \citet{siek2010threesomes}, \citet{garcia2013calculating} 
implement all other blame strategies with Threesome Calculus. 
He claims that 
coercion with labels is easy to understand but hard to implement, 
and that 
threesome with labels, however, is easy to implement but hard to understand.
His claim is later affirmed by the group of people who develop threesome 
(\citet{siek2015blame}).
The connection 
between his Threesome Calculi and \citet{siek2009exploring}'s Coercion 
Calculi are established by the fact that the former are derived from the latter.

\citet{siek2015blame} revisit the coercion-based approach. They revise the 
\lazyUD{} Coercion Calculus by \citet{siek2012interpretations}, prove its 
space-efficiency and that it bi-simulate the \lazyUD{} Blame Calculus.

Last year, \citet{kuhlenschmidt2018efficient} present Grift, a space-efficient 
and blame-tracking compiler for a gradually typed language of the same name. 
This implementation is based on the \lazyD{} Coercion Calculus by 
\citet{siek2012interpretations}.
Their result suggests that implementing coercion is practical. 
%They implement 
%tuples, which is the first time gradual product types is considered. Their 
%treatment on tuples, however, is shown incorrect 
%\cite{Gradual-TypingType-based-casts}.

@@

So far, it seems that the coercion-based approach is the best way to combine 
space-efficiency and blame-tracking: coercion is claimed easy to 
understand \cite{garcia2013calculating}\cite{siek2015blame} and is shown 
easy enough to implement in a compiler \cite{kuhlenschmidt2018efficient}.
Coercion is not satisfactory, however, in at least three aspects.
Firstly, its compose function is not structurally recursive, which makes 
it difficult to study coercion formally.
For instance, many developers of Grift report that it is tricky to convince 
their proof assistants that composition terminates. 
Secondly, it is desirable to understand the nature of ``normalized cast'', but 
the coercion-based approach capture this concept indirectly because normalized 
coercion is defined by taking a subset of coercion.
Last by not least, the memory representation of normal coercions is not very 
compact.

@@

Our hyper-coercion considers sum types and product types, which are not 
considered in all proofs above. Adding each of them requires us to add only one 
new constructor to a component of hyper-coercion. This suggests that 
hyper-coercion might scale better than super-coercion.

We prove formally that the \lazyD{} (resp. \lazyUD{}) hyper-coercion calculus 
bi-simulates the \lazyD{} (resp. \lazyUD{}) Blame Calculus. This is possibly 
the first bi-simulation proof for the \lazyD{} Blame Calculus.

@@

\begin{figure}
	Syntax
	\[
	\begin{array}{rclr}
	\stxrule{T}{types}{
		\star \mid{}
		P
	}
	\stxrule{P}{pre-types}{
		\POOunit \mid
		\POOfun{T_1}{T_2} \mid
		\POOprod{T_1}{T_2} \mid
		\POOsum{T_1}{T_2}
	}
	\stxrule{e}{terms}{
		\eOOvar{x} \mid{}
		\eOOsole{} \mid{}
		\eOOlam{T_1}{T_2}{x}{e} \mid
		\eOOapp{e_1}{e_2}
	}
	\stxrulecont{
		\eOOcons{e_1}{e_2} \mid
		\eOOcar{e} \mid
		\eOOcdr{e}
	}
	\stxrulecont{
		\eOOinl{e} \mid
		\eOOinr{e} \mid
		\eOOcase{e_1}{e_2}{e_3}
	}
	\stxrulecont{
		\eOOcast{e}{T_1}{l}{T_2} \mid
		\eOOblame{l}
	}
	\stxrule{o}{observations}{
		\oOOinj \mid
		\oOOsole \mid
		\oOOfun \mid
		\oOOcons \mid
		\oOOinl \mid
		\oOOinr \mid
		\oOOblame{l}
	}
	\end{array}
	\]

	Consistency
	\fbox{$ T_1 \sim T_2 $}
	\begin{gather*}
	\inference{}{
		\star \sim \star
	} \quad
	\inference{}{
		\star \sim P
	} \quad
	\inference{}{
		P \sim \star
	} \\
	\inference{}{
		\iota \sim \iota
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \rightarrow T_1 \sim S_2 \rightarrow T_2
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \times T_1 \sim S_2 \times T_2
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \plus T_1 \sim S_2 \plus T_2
	}
	\end{gather*}

	Shallow-consistency
	\fbox{$ T_1 \smile T_2 $}
	\begin{gather*}
	\inference{}{
		\star \smile \star
	} \quad
	\inference{}{
		\star \smile P
	} \quad
	\inference{}{
		P \smile \star
	} \\
	\inference{}{
		\iota \smile \iota
	} \quad
	\inference{}{
		T_{11} \rightarrow T_{12} \smile T_{21} \rightarrow T_{22}
	} \quad
	\inference{}{
		T_{11} \times T_{12} \smile T_{21} \times T_{22}
	} \quad
	\inference{}{
	T_{11} \plus T_1 \smile S_2 \plus T_2
	}
	\end{gather*}

	Term typing
	\fbox{$ \Gamma \vdash e \vdash T $}
	\begin{gather*}
		\inference{
			S \sim T & \Gamma \vdash e : S 
		}{
			\judgetype{\Gamma}{\eOOcast{e}{T_1}{l}{T_2}}{T_2}
		} \quad
		\inference{
		}{
			\judgetype{\Gamma}{\eOOblame{l}}{T}
		}
	\end{gather*}

@@

%$ S \sim T $ reads ``$ S $ and $ T $ are consistent''.
%Two types are consistent if one of them is $ \star $, or they have the same 
%top-most type constructor and the corresponding sub-parts are consistent. 
%The intuition of $ S \sim T $ is that $ S $ and $ T $ have no conflict type 
%information. Consistency is reflexive and symmetric, but not transitive.

%One might be tempted to conclude that inconsistency is the root of all runtime 
%type errors (blames). In the setting of gradual typing, however, runtime 
%type-checking is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
top-most type constructor. Shallow-inconsistency is the root of all blames in 
all lazy strategies -- casting a value to a shallowly inconsistent type leads 
to a blame immediately. Shallow-consistency is also reflexive, symmetric, and 
not transitive.

@@

\begin{figure}
	Syntax
	\[
	\begin{array}{rclr}

	\stxrule{v}{values}{
		\vOOfun{\rho}{x}{b} \mid
		\vOOtt{} \mid
		\vOOcons{v_1}{v_2} \mid
		\vOOinl{v} \mid
		\vOOinr{v} \mid		
		\vOOcast{v}{c}
	}
	\stxrule{c}{casts}{
		\cOOcast{T_1}{l}{T_2}
	}
	\stxrule{r}{cast results}{
		\rOOsucc{v} \mid
		\rOOfail{l}
	}
	\stxrule{s}{states}{
		\sOOinspect{e}{\rho}{\kappa} \mid{}
		\sOOreturn{v}{\kappa} \mid{}
		\sOOhalt{o}
	}

	\stxrule{\kappa}{continuations}{
		\mathtt{mt} \mid{}
		\mathtt{cons_1} \; e \; \rho \; \kappa \mid{}
		\mathtt{cons_2} \; v \; \kappa \mid{}
		\mathtt{inl} \; \kappa \mid{}
		\mathtt{inr} \; \kappa
	}
	\stxrulecont{
		\mathtt{app_1} \; e \; \rho \; \kappa \mid{}
		\mathtt{app_2} \; v \; \kappa \mid{}
		\mathtt{car} \; \kappa \mid{}
		\mathtt{cdr} \; \kappa \mid
		\mathtt{case_1} \; e_1 \; e_2 \; \rho \; \kappa
	}
	\stxrulecont{	
		\mathtt{case_2} \; v   \; e   \; \rho \; \kappa \mid{}
		\mathtt{case_3} \; v_1 \; v_2 \; \rho \; \kappa \mid
		\langle c \rangle \kappa
	}
	\end{array}
	\]

	Value typing \fbox{$ v : T $}
	\begin{gather*}
	\dots \quad
	\inference{
		v : P &
		c : P \Longrightarrow T &
		P \smile T
	}{
		\vOOcast{v}{\cOOcast{P}{l}{T}} : T
	}
	\end{gather*}

	Continuation typing \fbox{$ \kappa : T_1 \Longrightarrow T_2 $}
	\begin{gather*}
	\dots \quad
	\inference{
		c : T_1 \Longrightarrow T_2 &
		\kappa : T_2 \Longrightarrow T_3
	}{
		\langle c \rangle \kappa : T_1 \Longrightarrow T_3
	}
	\end{gather*}

	Reduction \fbox{$ s_1 \longrightarrow_B s_2 $}
	\[
	\begin{array}{rclr}

	& \vdots \\
	\redrule{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{\langle\cOOcast{T_1}{l}{T_2}\rangle\kappa}
	}{}
	\redrule{
		\sOOreturn{v_1}{(\mathtt{app_2} \; \vOOcast{v_2}{
				\cOOcast{\POOfun{T_{11}}{T_{12}}}{l}{\POOfun{T_{21}}{T_{22}}}
			} \; \kappa)}
	}{
		\sOOreturn{v_1}{
		\langle\cOOcast{T_{21}}{l}{T_{11}}\rangle
		(\mathtt{app_2} \; v_2 \; 
		\langle\cOOcast{T_{12}}{l}{T_{22}}\rangle \kappa)}
	}{}

%	\redrule{
%		\sOOreturn{v}{(
%			\mathtt{app_2} \;
%			(\vOOfun{\rho}{x}{e}) \;
%			\kappa
%		)}
%	}{\sOOinspect{e}{\rho{}[x := v]}{\kappa}}{}

	\redrule{
		\sOOreturn{
			\vOOcast{v}{\cOOcast{\POOprod{T_{11}}{T_{12}}}{l}{
					\POOprod{T_{21}}{T_{22}}}}
			}{(\mathtt{car} \; \kappa)}
	}{
		\sOOreturn{v}{(
			\mathtt{car} \;
			\langle \cOOcast{T_{11}}{l}{T_{21}} \rangle \kappa
			)}
	}{}

	\redrule{
		\sOOreturn{
			\vOOcast{v}{\cOOcast{\POOprod{T_{11}}{T_{12}}}{l}{
					\POOprod{T_{21}}{T_{22}}}}
		}{(\mathtt{cdr} \; \kappa)}
	}{
		\sOOreturn{v}{(
			\mathtt{cdr} \;
			\langle \cOOcast{T_{12}}{l}{T_{22}} \rangle \kappa
			)}
	}{}

	\redrule{
		\sOOreturn{v_3}{(
			\mathtt{case_3} \;
			v_1 \; v_2 \;
			\kappa
		)}
	}{case(v_1,
	\mathtt{app_2} \; v_2 \; \kappa,
	\mathtt{app_2} \; v_3 \; \kappa
	)}{}

	\redrule{
		\sOOreturn{v}{(
			\mathtt{cast} \; c \; \kappa
		)}
	}{
		\sOOreturn{u}{\kappa}
		\\ &
	}{
		\text{where} \; applyCast(c,v) = \rOOsucc{u}
	}

	\redrule{
		\sOOreturn{v}{(
			\mathtt{cast} \; c \; \kappa
			)}
	}{
		\sOOhalt{(\oOOblame{l})}
		\\ &
	}{
		\text{where} \; applyCast(c,v) = \rOOfail{l}
	}
	\end{array}
	\]

	Case
	\[
	\begin{array}{rclr}
	\funrule{case(\vOOinl{v},\kappa_1,\kappa_2)}{
		\sOOreturn{v}{\kappa_1}
	}{}
	\funrule{case(\vOOinr{v},\kappa_1,\kappa_2)}{
		\sOOreturn{v}{\kappa_2}
	}{}
	\funrule{case(\vOOcast{v}{\cOOcast{\POOsum{T_{11}}{T_{12}}}{l}{\POOsum{T_{21}}{T_{22}}}},\kappa_1,\kappa_2)}{
		case(v,
		\langle \cOOcast{T_{11}}{l}{T_{21}} \rangle \kappa_1,
		\langle \cOOcast{T_{12}}{l}{T_{22}} \rangle \kappa_2)
	}{}
	\end{array}
	\]

	Apply cast
	\[
	\begin{array}{rclr}
	\funrule{
		applyCast(\cOOcast{\star}{l}{\star},v)
	}{
		\rOOsucc{v}
	}{}
	\funrule{
		applyCast(\cOOcast{\star}{l}{P_2},\vOOcast{v}{\cOOcast{P_1}{l'}{\star}})
	}{
		applyCast(\cOOcast{P_1}{l}{P_2},v)
	}{}
	\funrule{
		applyCast(\cOOcast{P}{l}{\star},v)
	}{
		\rOOsucc{\vOOcast{v}{\cOOcast{P}{l}{\star}}}
	}{}
	\funrule{
		applyCast(\cOOcast{P_1}{l}{P_2},v)
	}{
		\rOOsucc{\vOOcast{v}{\cOOcast{P_1}{l}{P_2}}}
	}{P_1 \smile P_2}
	\funrule{
		applyCast(\cOOcast{P_1}{l}{P_2},v)
	}{
		\rOOfail{l}
	}{\neg P_1 \smile P_2}

	\end{array}
	\]

	\caption{Dynamic semantics of the \lazyD{} Blame Calculus}
	\label{ld-blame-cek}
\end{figure}

Fig.~\ref{ld-blame-cek} defines the dynamic semantcis of the \lazyD{} Blame 
Calculus in the style of CEK machines (\citet{felleisen1986control}). 
We choose CEK machine because to achieve space efficiency, we need to change 
continuations. Thus, abstract machines that include continuation in states is 
quite convenient for us. And CEK is one of the simplest machine of this kind.

Let $ c $ ranges over casts. A cast is a triple of a type, a label, and a type.

Let $ v $ ranges over values. A value is either a function, the 
element of unit, a pair, a left injection, a right injection, or a casted value.
The cast around a casted value is not arbitrary, as shown by value typing: the 
source type must be a pre-type, and the source and the target must be 
shallow-consistent.

Let $ r $ ranges over cast results. A cast result is either a success, which 
brings a value, or a failure, which brings a blame label.

Let $ s $ ranges over machine states. A state is either looking at an 
expression to decide what to do next, returning a value to a continuation, or 
halting with an observation.

Let $ \kappa $ ranges over continuations. $ \mathtt{mt} $ is the top 
continuation. The following continuations correspond to expressions. For 
example, $ (\mathtt{cons_1} \; e \; \rho \; \kappa) $ is the continuation where 
we are waiting for the value of the first argument to a $ \mathtt{cons} $. And 
the last continuation, $ \langle c \rangle \kappa $ is to cast the value before 
returning to $ \kappa $.

In the reduction rules, to evaluate a cast expression, the machine move the 
cast to the continuation and evaluate the inner expression. To apply a casted 
function, the machine firstly cast $ v_1 $, the operand, then apply the casted 
operand to $ v_2 $, the un-casted function, and finally cast the result of the 
function application. To take out the first (resp. second) part of a casted 
pair, the machine firstly take out the first (resp. second) part of $ v $, the 
un-casted pair, and cast the result. To case-split a sum, the machine looks 
deep inside the value to find out whether it is a left injection or a right 
one. Along the way, the machine accumulate the pending casts onto the 
continuations. To cast a value, the machine try to apply the cast to the value. 
If the cast succeeds, the machine returns the result to the next continuation. 
If the cast fails, the machine halts with the blame label.

@@

	\fbox{$ m \fatsemi^\ell (m,t) = (m,t) $}
	\[ 
	\begin{array}{rclclr}
	\comprule{U}{(U,t)}{
		(U,t)
	}{}
	\comprule{c_1 \rightarrow c_2}{(c_3 \rightarrow c_4,t)}{
		(c_3 \fatsemi^{\ell} c_1 \rightarrow c_2 \fatsemi^\ell c_4, t)
	}{}
	\comprule{c_1 \times c_2}{(c_3 \times c_4,t)}{
		(c_1 \fatsemi^{\ell} c_3 \times c_2 \fatsemi^\ell c_4, t)
	}{}
	\comprule{c_1 \plus c_2}{(c_3 \plus c_4,t)}{
		(c_1 \fatsemi^{\ell} c_3 \plus c_2 \fatsemi^\ell c_4, t)
	}{}
	\comprule{b_1}{(b_2,t)}{
		(b_1,\bot^l)
	}{
		\ell = l \; \text{and} \;
		\neg b_1 \smile b_2 
	}
	\end{array}
	\]

@@

	\fbox{$ id( P ) = m $}
	\[
	\begin{array}{rclr}
	\funrule{id(\POOunit)}{\POOunit}{}
	\funrule{id(\POOfun{T_1}{T_2})}{
		\POOfun{id(T_1)}{id(T_2)}
	}{}
	\funrule{id(\POOprod{T_1}{T_2})}{
		\POOprod{id(T_1)}{id(T_2)}
	}{}
	\funrule{id(\POOsum{T_1}{T_2})}{
		\POOsum{id(T_1)}{id(T_2)}
	}{}
	\end{array}
	\]
